```{r load data}
setwd('/media/jidde/Western Dig/analysis_backup_18_08_2021/levodopa project')
data <- readRDS("data_levodopa_29_01_21.rds")
```


Load Libraries
```{r analysis}
library(ggplot2)
library(dplyr)
library(lme4)
library(car)
library(lmerTest)
library(visreg)
```


# Statistic analysis

Write Cohens d function
```{r cohensDfunction}
lme.dscore<-function(mod,data,type){
  if (type=="lme4") {
    mod1<-lmerTest::lmer(mod,data=data)
    eff<-cbind(summary(mod1)$coefficients[,4],summary(mod1)$coefficients[,3])
  }
 
  if (type=="nlme") {
    eff=cbind(summary(mod)$tTable[,4],summary(mod)$fixDF$terms)
  }
 
  colnames(eff)<-c("t","df")
  eff<-as.data.frame(eff)
  eff$d<-(2*eff$t)/sqrt(eff$df)
  eff<-eff[-1,]
  return(eff)
}
```

# Models for VAI

Model with Gender as fixed effect & random intercepts for subjects
```{r}
m1 <- lmer(VAI ~ Gender + (1|participant.id), data = data)
summary(m1)
```


Model with group and Gender as fixed & random intercepts for subjects
```{r}
summary(m2 <- lmer(VAI ~ group + Gender + (1|participant.id), data = data))
anova(m1, m2)
anova(m2)
```
No improvement.

Add language as fixed effect
```{r}
summary(m3 <- lmer(VAI ~ group + Gender + lang + (1|participant.id), data = data))
anova(m2, m3)
```
No improvement.

Add time of day as fixed effect.
```{r}
summary(m4 <- lmer(VAI ~ group + Gender + time + (1|participant.id), data = data))
anova(m2, m4)
```
No improvement.

Add day as fixed effect.
```{r}
summary(m5 <- lmer(VAI ~ group + Gender + day + (1|participant.id), data = data))
anova(m2, m5)
anova(m5)
```
No improvement.

Add age as fixed effect.
```{r}
summary(m6 <- lmer(VAI ~ group + Gender + age + (1|participant.id), data = data))
anova(m2, m6)
```
No improvement.

## Interactions
Interaction between group and Gender.
```{r}
summary(m7 <- lmer(VAI ~ group*Gender + (1|participant.id), data = data))
anova(m7, m2)
```
Sig. improv, but not sign. when fitted with REML.

Interaction between group and language.
```{r}
summary(m8 <- lmer(VAI ~ group*lang + Gender + (1|participant.id), data = data))
anova(m8, m2)
```
No improvement. 

Interaction between group and time.
```{r}
summary(m9 <- lmer(VAI ~ group*time + Gender + (1|participant.id), data = data))
anova(m9, m2)
```
No improvement.

Interaction between language and time.
```{r}
summary(m10 <- lmer(VAI ~ group + Gender + lang*time + (1|participant.id), data = data))
anova(m2, m10)
```
No improvement.

Interaction between Gender and lang.
```{r}
summary(m11 <- lmer(VAI ~ group + Gender*lang + (1|participant.id), data = data))
anova(m2, m11)
```
No improvement.

## Random slopes

```{r}
m3_randslo_group <- lmer(VAI ~ group + Gender + (1+session|participant.id), data = data,  control = lmerControl(optimizer = "bobyqa")) # fails to converge
m3_randslo_day <- lmer(VAI ~ group + Gender + (1+day|participant.id), data = data,  control = lmerControl(optimizer = "bobyqa")) # boundary (singular) fit
```
Fails to converge.


# Check assumptions final model
```{r}
lme.dscore(m2, data, type="lme4")

acf(resid(m2)) # autocorrelation seems ok

qqnorm(resid(m2)) 
qqline(resid(m2)) # looks OKish

plot(fitted(m2), resid(m2)) 
```




### MODELS FOR VAI PD ONLY 
Model with only PD, state as fixed effect & random intercepts for subjects
```{r}
summary(m1PD <- lmer(VAI ~ state + (1|participant.id), data = data[data$group=="PD",]))
```

Model with only PD, session as fixed effect & random intercepts for subjects
```{r}
summary(m2PD <- lmer(VAI ~ session + (1|participant.id), data = data[data$group=="PD",]))
anova(m1PD, m2PD)
anova(m2PD)
```
Stick with state.


Model with state and Gender as fixed effect & random intercepts for subjects
```{r}
summary(m3PD <- lmer(VAI ~ state + Gender + (1|participant.id), data = data[data$group=="PD",]))
anova(m1PD, m3PD)
anova(m3PD)
```
Model 3PD has sign. lower AIC

Now we add language to the model.
```{r}
summary(m4PD <- lmer(VAI ~ state + Gender + lang + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m4PD)
```
No improvement.

Add day to the model.
```{r}
summary(m5PD <- lmer(VAI ~ state + Gender + day + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m5PD)
anova(m5PD)
```
No improvement.

Add time of day to the model.
```{r}
summary(m6PD <- lmer(VAI ~ state + Gender + time + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m6PD)
```
No improvement.

Add levodopa
```{r}
summary(m7PD <- lmer(VAI ~ state + Gender + levodopa + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD,m7PD)
```
No improvement.

Add diagnosis
```{r}
summary(m8PD <- lmer(VAI ~ state + Gender + diagnosis + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD,m8PD)
summary(m8PD)
```
Sign improvement in model comparison, but no sign effect of predictor when fitted with REML (p=0.09).

Add age
```{r}
summary(m9PD <- lmer(VAI ~ state + Gender + age + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD,m9PD)
```
No improvement.

## Interactions

Add state*Gender
```{r}
summary(m10PD <- lmer(VAI ~ state*Gender + (1|participant.id), data = data[data$group=="PD",]))
anova(m10PD, m3PD)
```
No improvement.

Add state*language
```{r}
summary(m11PD <- lmer(VAI ~ state*lang + Gender + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m11PD)
```
No improvement.

Add state*day interaction
```{r}
summary(m12PD <- lmer(VAI ~ state*day + Gender + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m12PD)
anova(m12PD)
```
No added value.


Add lang*time interaction
```{r}
summary(m13PD <- lmer(VAI ~ state + Gender + lang*time + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m13PD)
```
No improvement.

Add gender*language interaction
```{r}
summary(m14PD <- lmer(VAI ~ state + Gender*lang + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m14PD)
summary(m14PD)
```
Significant, but not reliable since there was only one female in the Dutch sample.

Add age*language interaction
```{r}
m15PD <- lmer(VAI ~ state + Gender + lang*age + (1|participant.id), data = data[data$group=="PD",])
anova(m3PD, m15PD)
```
No improvement.


Add state*time of day interaction
```{r}
summary(m16PD <- lmer(VAI ~ state*time + Gender + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m16PD)
```
No improvement.

Add language*day interaction
```{r}
summary(m17PD <- lmer(VAI ~ state*day + Gender + (1|participant.id), data = data[data$group=="PD",]))
anova(m3PD, m17PD)
anova(m17PD)
```


## Random slopes

```{r}
m3PDrandslo_state <- lmer(VAI ~ state + Gender + (1+state|participant.id), data = data[data$group=="PD",]) 
m3PDrandslo_time <- lmer(VAI ~ state + Gender  + (1+time|participant.id), data = data[data$group=="PD",])
m3PDrandslo_session <- lmer(VAI ~ state + Gender  + (1+session|participant.id), data = data[data$group=="PD",]) # singular fit

anova(m3PD, m3PDrandslo_state)
anova(m3PD, m3PDrandslo_time)

```
No improvements.

# Check assumptions final model
```{r}
lme.dscore(m3PD, data, type="lme4")

acf(resid(m3PD)) # autocorrelation seems ok

qqnorm(resid(m3PD)) 
qqline(resid(m3PD)) # looks OK

plot(fitted(m3PD), resid(m3PD)) # looks OKish
```

### Fatigue
```{r}
m1PDf <- lmer(VAI ~ state + (1|participant.id), data = data[data$group=="PD" & data$lang == "Slovenian",])
```


```{r}
m2PDf <- lmer(VAI ~ state + fatigue + (1|participant.id), data = data[data$group=="PD" & data$lang == "Slovenian",])
anova(m1PDf, m2PDf)
```


### plots
```{r}

ggplot(data=data, aes(x=participant.id, y=VSA, fill=group)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) +   #stat_summary(aes(label=group), fun.y=mean, geom="text", size=2) + 
  #facet_wrap(Gender~lang, nrow = 1, scales = "free") + 
  facet_wrap(Gender~lang, scales = "free")+
  labs(x = "Participant") +
  theme_bw()

ggplot(data=data, aes(x=participant.id, y=VAI, fill=group)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) +   #stat_summary(aes(label=group), fun.y=mean, geom="text", size=2) + 
  #facet_wrap(Gender~lang, nrow = 1, scales = "free") + 
  facet_wrap(Gender~lang, scales = "free")+
  labs(x = "Participant") +
  theme_bw()


ggplot(data=data[data$group=="PD",], aes(x=participant.id, y=VSA, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + labs(x = "participant") 


ggplot(data=data[data$group=="PD",], aes(x=participant.id, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "participant") + theme_bw()




#example of individual variation
ggplot(data=data[data$participant.id=="NL02",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 0.95)) + labs(x = "Recording day")

#example of individual variation
ggplot(data=data[data$participant.id=="NL02",], aes(x=day, y=VSA, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + labs(x = "Recording day", y = "VSA")


ggplot(data, aes(x=group, y=VSA, fill=Gender)) + geom_boxplot() + labs(y = "VSA")  +
  theme_bw()

```

```{r plots per participant}

### per participant ###
par(mfrow=c(4,2))

#example of individual variation
ggplot(data=data[data$participant.id=="NL01",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")


#example of individual variation
ggplot(data=data[data$participant.id=="NL03",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")

#example of individual variation
ggplot(data=data[data$participant.id=="NL04",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")

#example of individual variation
ggplot(data=data[data$participant.id=="NL05",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")


#example of individual variation
ggplot(data=data[data$participant.id=="NL06",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")

#example of individual variation
ggplot(data=data[data$participant.id=="NL07",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")
```

Figures to report:
```{r}
#example of individual variation
ggplot(data=data[data$participant.id=="NL02",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 0.925)) + labs(x = "Recording day") + theme_bw()


#example of individual variation
ggplot(data=data[data$participant.id=="NL02",], aes(x=day, y=VSA, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(100000, 225000)) + labs(x = "Recording day") + theme_bw()

```

# Session info
```{r}
sessionInfo()
```

