```{r load data}
setwd('/media/jidde/Western Dig/analysis_backup_18_08_2021/levodopa project')
data <- readRDS("data_levodopa_29_01_21.rds")
```


Load Libraries
```{r analysis}
library(ggplot2)
library(plyr)
library(dplyr)
library(lme4)
library(car)
library(lmerTest)
library(visreg)
```


# Statistic analysis

Write Cohens d function
```{r cohensDfunction}
lme.dscore<-function(mod,data,type){
  if (type=="lme4") {
    mod1<-lmerTest::lmer(mod,data=data)
    eff<-cbind(summary(mod1)$coefficients[,4],summary(mod1)$coefficients[,3])
  }
 
  if (type=="nlme") {
    eff=cbind(summary(mod)$tTable[,4],summary(mod)$fixDF$terms)
  }
 
  colnames(eff)<-c("t","df")
  eff<-as.data.frame(eff)
  eff$d<-(2*eff$t)/sqrt(eff$df)
  eff<-eff[-1,]
  return(eff)
}
```

# Models for VSA

Base model
```{r}
summary(m0 <- lmer(VSA ~ (1|participant.id), data = data))
summary(m0)
```

Model with gender as fixed effect & random intercepts for subjects
```{r}
summary(m1 <- lmer(VSA ~ Gender + (1|participant.id), data = data))
anova(m0, m1)
visreg(m1)
```


Model with group and Gender as fixed & random intercepts for subjects
```{r}
summary(m2 <- lmer(VSA ~ group + Gender + (1|participant.id), data = data))
anova(m1, m2)
anova(m2)
```
Model 2 has sign. lower AIC

Add language as fixed effect
```{r}
summary(m3 <- lmer(VSA ~ group + Gender + lang + (1|participant.id), data = data))
anova(m2, m3)
```
No improvement.

Add time of day as fixed effect.
```{r}
summary(m4 <- lmer(VSA ~ group + Gender + time + (1|participant.id), data = data))
anova(m2, m4)
```
Time is no sign. predictor, stick with model3.

Add day as fixed effect.
```{r}
summary(m5 <- lmer(VSA ~ group + Gender + day + (1|participant.id), data = data))
anova(m2, m5)
anova(m5)
```
No improvement.

Add age as fixed effect.
```{r}
summary(m6 <- lmer(VSA ~ group + Gender + age + (1|participant.id), data = data))
anova(m2, m6)
```
No improvement.

## Interactions
Interaction between group and Gender.
```{r}
summary(m7 <- lmer(VSA ~ group*Gender + (1|participant.id), data = data))
anova(m7, m2)
```
No improvement.

Interaction between group and language.
```{r}
summary(m8 <- lmer(VSA ~ group*lang + Gender + (1|participant.id), data = data))
anova(m8, m2)
```
No improvement. 

Interaction between group and time.
```{r}
summary(m9 <- lmer(VSA ~ group*time + Gender + (1|participant.id), data = data))
anova(m9, m2)
```
No improvement.

Interaction between language and time.
```{r}
summary(m10 <- lmer(VSA ~ group + Gender + lang*time + (1|participant.id), data = data))
anova(m2, m10)
```
No improvement.


Interaction between Gender and lang.
```{r}
summary(m11 <- lmer(VSA ~ group + Gender*lang + (1|participant.id), data = data))
anova(m2, m11)
```
No improvement.

Interaction between Gender and age.
```{r}
summary(m12 <- lmer(VSA ~ group + Gender*age + (1|participant.id), data = data))
anova(m2, m12)
```
No improvement. 

## Random slopes

```{r}
m3_randslo_group <- lmer(VSA ~ group + Gender + (1+session|participant.id), data = data,  control = lmerControl(optimizer = "bobyqa")) # fails to converge
m3_randslo_day <- lmer(VSA ~ group + Gender + (1+day|participant.id), data = data,  control = lmerControl(optimizer = "bobyqa")) # boundary (singular) fit
```
No improvement.


# Check assumptions final model
```{r}
lme.dscore(m2, data, type="lme4")

acf(resid(m2)) # autocorrelation seems ok

qqnorm(resid(m2)) 
qqline(resid(m2)) # looks OKish

plot(fitted(m2), resid(m2)) 
```
Looks fine.



### MODELS FOR VSA PD ONLY

Create new df with only PD:
```{r}
dataPD <- subset(data, group == "PD")
```


Model with only PD, state as fixed effect & random intercepts for subjects
```{r}
summary(m1PD <- lmer(VSA ~ state + (1|participant.id), data = dataPD))
summary(m1PD)
```
No improvement.


Model with only PD, session as fixed effect & random intercepts for subjects
```{r}
summary(m2PD <- lmer(VSA ~ session + (1|participant.id), data = dataPD))
anova(m1PD, m2PD)
anova(m2PD)
```
Stick with state.


Model with state and Gender as fixed effect & random intercepts for subjects
```{r}
summary(m3PD <- lmer(VSA ~ state + Gender + (1|participant.id), data = dataPD))
anova(m1PD, m3PD)
anova(m3PD)
```
Model 3PD has sign. lower AIC

Now we add language to the model.
```{r}
summary(m4PD <- lmer(VSA ~ state + Gender + lang + (1|participant.id), data = dataPD))
anova(m3PD, m4PD)
```
The model including language has a sign. lower AIC, but effect is not sign. when fitted with REML. 

Add day to the model.
```{r}
summary(m5PD <- lmer(VSA ~ state + Gender + day + (1|participant.id), data = dataPD))
anova(m3PD, m5PD)
anova(m5PD)
```
Day does not improve the fit.

Add time of day to the model.
```{r}
summary(m6PD <- lmer(VSA ~ state + Gender + time + (1|participant.id), data = dataPD))
anova(m3PD, m6PD)
```
No improvement.

Add levodopa
```{r}
summary(m7PD <- lmer(VSA ~ state + Gender + levodopa + (1|participant.id), data = dataPD))
anova(m3PD,m7PD)
```
No improvement.

Add diagnosis
```{r}
summary(m8PD <- lmer(VSA ~ state + Gender + diagnosis + (1|participant.id), data = dataPD))
anova(m3PD,m8PD)
summary(m8PD)
```
Sign improvement in model comparison, but no sign effect of predictor when fitted with REML (p=0.08).

Add age
```{r}
summary(m9PD <- lmer(VSA ~ state + Gender + age + (1|participant.id), data = dataPD))
anova(m3PD,m9PD)
```
No improvement.

## Interactions

Add state*Gender
```{r}
summary(m10PD <- lmer(VSA ~ state*Gender + (1|participant.id), data = dataPD))
anova(m10PD, m3PD)
```
No change.

Add state*language
```{r}
summary(m11PD <- lmer(VSA ~ state*lang + Gender + (1|participant.id), data = dataPD))
anova(m3PD, m11PD)
summary(m11PD)
```
Sign in model comparison using ML, but effect is not sign, when using REML (p=0.06).

Add state*day interaction
```{r}
summary(m12PD <- lmer(VSA ~ state*day + Gender + (1|participant.id), data = dataPD))
anova(m3PD, m12PD)
anova(m12PD)
```
No added value.


Add lang*time interaction
```{r}
summary(m13PD <- lmer(VSA ~ state + Gender + lang*time + (1|participant.id), data = dataPD))
anova(m3PD, m13PD)
```
No improvement.

Add Gender*language interaction
```{r}
data$state <- as.factor(data$state)
summary(m14PD <- lmer(VSA ~ state + Gender*lang + (1|participant.id), data = dataPD))
visreg(m14PD, "state", "Gender", xlab="state", ylab="VSA", gg=TRUE, cond=list(lang = "Dutch")) + ggtitle("Dutch") + ylim(0,600000) + theme_bw() 
visreg(m14PD, "state", "Gender", xlab="state", ylab="VSA", gg=TRUE, ylim=c(0,6000000), cond=list(lang = "Slovenian")) + ylim(0,600000) + ggtitle("Slovenian")+ theme_bw()
anova(m3PD, m14PD)
anova(m14PD)
```
Significant, but not reliable since there was only one female in the Dutch sample.


Add language*age interaction
```{r}
summary(m15PD <- lmer(VSA ~ state + Gender + lang*age + (1|participant.id), data = dataPD))
anova(m3PD, m15PD)
summary(m15PD)
```
Sign. after model comparison but effect itself is not sign when fitted with REML.


Add state*time interaction
```{r}
summary(m16PD <- lmer(VSA ~ state*time + Gender + (1|participant.id), data = dataPD))
anova(m3PD, m16PD)
```
No improvement.


Add lang*day interaction
```{r}
summary(m17PD <- lmer(VSA ~ state + Gender + lang*day + (1|participant.id), data = dataPD))
anova(m3PD, m17PD)
anova(m17PD)
```
No improvement.


## Random slopes

```{r}
summary(m3PDrandslo_state <- lmer(VSA ~ state + Gender + (1+state|participant.id), data = dataPD)) # singular fit
summary(m3PDrandslo_time <- lmer(VSA ~ state + Gender + (1+time|participant.id), data = dataPD)) # singular fit
summary(m3PDrandslo_session <- lmer(VSA ~ state + Gender + (1+session|participant.id), data = dataPD)) # singular fit
```
Models become too complex.

# Check assumptions final model
```{r}
lme.dscore(m14PD, data, type="lme4")

acf(resid(m14PD)) # autocorrelation seems ok

qqnorm(resid(m14PD)) 
qqline(resid(m14PD)) # looks OK

plot(fitted(m14PD), resid(m14PD)) 
```


### Fatigue
```{r}
summary(m1PDf <- lmer(VSA ~ state + (1|participant.id), data = data[data$group=="PD" & data$lang == "Slovenian",]))
```



```{r}
summary(m2PDf <- lmer(VSA ~ state + fatigue + (1|participant.id), data = data[data$group=="PD" & data$lang == "Slovenian",]))
anova(m1PDf, m2PDf)
```
No improvement.


### plots
```{r}
data$group <- revalue(data$group, c("CTRL" = "Typical"))
         
         ggplot(data=data, aes(x=participant.id, y=VSA, fill=group)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) +
  facet_wrap(Gender~lang, nrow = 2, scales = "free_x") + labs(x = "participant")  +
  ggsci::scale_fill_npg() + theme_bw()

ggplot(data=data, aes(x=participant.id, y=VAI, fill=Gender)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + coord_cartesian(ylim = c(0.8, 1.3)) +
  facet_grid(lang~group,space="fixed",scales = "free_x")


ggplot(data=dataPD, aes(x=participant.id, y=VSA, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + labs(x = "participant") + labs(colour = "legend title") + theme_bw()


ggplot(data=dataPD, aes(x=participant.id, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "participant") + theme_bw()




#example of individual variation
ggplot(data=data[data$participant.id=="NL02",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 0.95)) + labs(x = "Recording day")

#example of individual variation
ggplot(data=data[data$participant.id=="NL02",], aes(x=day, y=VSA, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + labs(x = "Recording day", y = "VSA")


ggplot(data, aes(x=group, y=VSA, fill=Gender)) + geom_boxplot() + labs(y = "VSA") + theme_bw() +
  ggsci::scale_fill_npg()

```

```{r plots per participant}

### per participant ###
par(mfrow=c(4,2))

#example of individual variation
ggplot(data=data[data$participant.id=="NL01",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")

#VAI
ggplot(data=data[data$participant.id=="NL02",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 0.93)) + labs(x = "recording day") + theme_bw()

#VSA
ggplot(data=data[data$participant.id=="NL02",], aes(x=day, y=VSA, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 250000)) + labs(x = "recording day") + theme_bw()

#example of individual variation
ggplot(data=data[data$participant.id=="NL03",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")

#example of individual variation
ggplot(data=data[data$participant.id=="NL04",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")

#example of individual variation
ggplot(data=data[data$participant.id=="NL05",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")


#example of individual variation
ggplot(data=data[data$participant.id=="NL06",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")

#example of individual variation
ggplot(data=data[data$participant.id=="NL07",], aes(x=day, y=VAI, fill=session)) +
  geom_bar(stat="summary", fun.y=mean, position='dodge') +
   stat_summary(fun.data = "mean_se", geom = 'errorbar', position=position_dodge(.9), width = 0.2) + scale_fill_manual(values = c("#ff0000","#008500", "#66cc00","#005483","#6fb5e9")) + coord_cartesian(ylim = c(0.8, 1.3)) + labs(x = "Recording day")
```


# Session info
```{r}
sessionInfo()
```
